---
title: "Practice Activity 9.1: Regression"
author: "Jasmine Cabrera"
format:
  html:
    self-contained: true
    code-tools: true
    code-fold: true
execute:
  echo: true
  error: false
  messages: false
  warning: false
---

## 1. Data

The dataset [`mystery_animal.csv`](https://atheobold.github.io/stat-331/9_simulation-regression/Practice%20Activity/mystery_animal.csv) contains the weights of a particular animal before and after a year of eating only roasted duck.

```{r}
library(tidyverse)
animal <- read_csv(here::here("mystery_animal.csv"))
```

## 2. Linear Regression

Fit a linear regression to determine if the duck diet is associated with the animal gaining weight, losing weight, or neither.

```{r}
animal_lm <- animal |>
  lm(weight_after ~ weight_before, data = _) |>
  print()
```

## 3. Residuals

Plot the residuals (observed value minus predicted value) of your linear model versus `weight_before`.

There are a few different ways to obtain your residuals. My favorite is the `augment()` function from the **broom** package. I like this option a lot because it gives you all of the information from your linear regression in a tidy [tibble]{.underline}! Each row has the explanatory and response values, the fitted / predicted value (`.fitted`), and the residual (`.resid`).

```{r}
animal_lm |>
  broom::augment() |>
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point()
```
