---
title: "Lab 9: Baby Names"
author: "Jasmine Cabrera"
format:
  html:
    self-contained: true
    code-tools: true
    code-fold: true
execute:
  echo: true
  error: false
  messages: false
  warning: false
---

## 1 The Data

Our dataset(s) in this lab concerns baby names and their popularity over time. At [this link](https://www.dropbox.com/sh/czm61yzmvzl6xfb/AABEFdpy02aNRhZ_vcxrVAQSa?dl=0), you can find the names for ALL 50 states, in separate datasets organized by first letter. For each year, and for each name with at least 50 recorded babies born, we are given the counts of how many babies had that name.

## 2 Is My Name Not Cool Anymore?

Let's take a look at how the name "Allison" has changed over time. As my name begins with "A", you should download the `StateNames_A.csv` dataset from the link above.

```{r load in data}
library(tidyverse)
library(here)
library(DT)
library(gt)
library(kableExtra)

# Imports data
names <- read_csv(here("StateNames_A.csv"))

# Interactive Data Table
DT::datatable(names)
```

## 3 Summarizing & Visualizing the Number of Allisons

1.  Make a summary table of the number of babies named "Allison" for each state and the sex of the baby. Specifically, the table should have the following qualities:

-   each state should be its own row

-   and each sex should have its own column

-   if there were no babies born for that combination of state & sex there should be a 0 (not an `NA`)

The dataset has a column titled `Gender`, which contains two values `"F"` and `"M"`, representing "Female" and "Male". The sex someone was assigned at birth is different from their gender identity [(definitions)](https://www.plannedparenthood.org/learn/gender-identity/sex-gender-identity). Thus, this variable should be renamed to `Sex` or `Sex at Birth`.

2.  You should have seen in the table above that "Allison" is a name given overwhelmingly to babies assigned "female" at birth. So, create a new data set named `allison_f` which contains **only** the babies assigned `F`emale at birth.

This should be the data set you use for the remainder of the assignment.

3.  Make a visualization showing how the popularity of the name "Allison" has changed over the years. To be clear, each year should have **one** observation--the total number of Allisons born that year.

    ```{r}
    allison <- names |>
      rename(Sex = Gender) |>
      filter(Name == "Allison") |>
      group_by(State, Sex) |>
      summarize(Total = sum(Count), 
                .groups = "drop") |>
      pivot_wider(names_from = Sex,
                  values_from = Total,
                  values_fill = 0) |>
      print()

    allison |>
      knitr::kable(format = "html",
                   col.names = c("State",
                                 "Female",
                                 "Male"),
                   caption = "Frequency of Allisons in Each State") |>
      kableExtra::kable_classic(html_font = "Arial") |>
      kableExtra::kable_styling(bootstrap_options = "striped")


    allison_f <- names |>
      rename(Sex = Gender) |>
      filter(Name == "Allison") 
    
    allison_f|>
      group_by(Year) |>
      summarize(Total = sum(Count)) |>
      ggplot(mapping = aes(x = Year,
                           y = Total)) +
      geom_line() +
      labs(y = NULL,
           title = "Total Number of Allisons Born Between 1997-2014") +
      scale_x_continuous(breaks = seq(1997, 2014, by = 1), 
                         guide =  guide_axis(n.dodge = 2)) +
      scale_y_continuous(breaks = seq(5000, 9000, by = 1000), 
                         limits = c(5000, 8000)) + 
      theme(plot.title = element_text(face = "bold",
                                      hjust = 0.5))
    allison_f
    ```

**Source:**

https://stackoverflow.com/questions/55948425/why-is-scale-y-continuous-not-working-here-no-error-message-given

https://www.datanovia.com/en/blog/ggplot-title-subtitle-and-caption/#:\~:text=Change%20the%20font%20appearance%20%28text%20size%2C%20color%20and,font%20face%20include%3A%20%E2%80%9Cplain%E2%80%9D%2C%20%E2%80%9Citalic%E2%80%9D%2C%20%E2%80%9Cbold%E2%80%9D%20and%20%E2%80%9Cbold.italic%E2%80%9D.

## 4 Modeling the Number of Allisons

4.  Fit a linear model with the year as the explanatory variable, and the number of Allisons as the response. Similar to #3, each year should have **one** observation--the total number of Allisons born that year.

5.  Write out the estimated regression equation.

    **Estimated Total Number of Allisons = 209689.8 -101.5(Year)**

    \*\*\***Note:** Year must be substituted with the actual Year (i.e. 1997, 2014, etc.)

6.  Plot the *residuals* of the model, that is, the actual values minus the predicted values. Comment on the residuals - do you see any patterns?

There is slight fanning in the residuals plot with the residuals forming a concave up parabola.

The `augment()` function from the **broom** package may be helpful!

7.  What do you conclude from this model? Is my name not cool anymore?

We can not conclude that the name "Allison" is decreasing in popularity over the years 1997 to 2014 for babies assigned "female" at birth from this model. This is due to the slight curvature in the residuals plot that suggests that we should not use a straight line to model this data.

```{r}
allison_lm <- allison_f |>
  group_by(Year) |>
  summarize(Total = sum(Count)) |>
  lm(Total ~ Year, data = _) |>
  print()

allison_lm |>
  broom::augment() |>
  ggplot(mapping = aes(y = .resid, x = .fitted)) +
  geom_point() + 
  labs(x = "Fitted",
       y = NULL,
       title = "Residual Plot") +
  scale_x_continuous(breaks = seq(5000, 7000, by = 500), 
                     limits = c(5000, 7000)) + 
  scale_y_continuous(breaks = seq(-1000, 1000, by = 500), 
                     limits = c(-1250, 1250)) + 
  theme(plot.title = element_text(face = "bold",
                                  hjust = 0.5))

```

## 5 Spelling by State

In middle school I was **so** upset with my parents for not naming me "Allyson". Past my pre-teen rebellion, I'm happy with my name and am impressed when baristas spell it "Allison" instead of "Alison". But I don't have it as bad as my good friend Allan!

8.  Narrow the A name dataset (downloaded previously) down to only male-assigned babies named "Allan", "Alan", or "Allen". Make a plot comparing the popularity of these names over time.

It looks like you want to filter for a vector of values. What tools have you learned which can help you accomplish this task?

9.  In California, Allan's spelling of his name is the least common of the three but perhaps it's not such an unusual name for his home state of Pennsylvania. Compute the total number of babies born with each spelling of "Allan" in the year 2000, in Pennsylvania and in California. Specifically, the table should have the following qualities:

-   each spelling should be its own column

-   each state should have its own row

-   a `0` (not an `NA`) should be used to represent locations where there were no instances of these names

10. Convert your total counts to overall percents. That is, what was the percent breakdown between the three spellings in CA? What about in PA?

In the year 2000, 21.364% of all Male birthed babies were named some form of "Allan", "Alan", or "Allen" were born in California and 2.869% were born in Pennsylvania. Looking closer, in the California selection: 19.86% were named Allen, 65.34% were named Alan, and 14.785% were named Allan. In the Pennsylvania selection: 47.05% were named Allen, 42.85% were named Alan, and 1.01% were named Allan.

```{r allan}
allans <- tibble(Name = c("Allan", 
                          "Alan", 
                          "Allen"))

allan_names <- names |>
  filter(Gender == "M") |>
  semi_join(allans, by = "Name") |>
  rename("Sex" = "Gender") 

allan_names |>
  group_by(Name, Year) |>
  summarize(Total = sum(Count)) |>
  ggplot(mapping = aes(x = Year,
                       y = Total,
                       color = Name)) +
  geom_line() +
  labs(y = NULL,
       title = "Total Number of Allans Born Between 1997-2014") +
  scale_x_continuous(breaks = seq(1997, 2014, by = 1), 
                     guide =  guide_axis(n.dodge = 2)) +
  scale_y_continuous(breaks = seq(0, 3500, by = 1000), 
                     limits = c(0, 3500)) + 
  theme(plot.title = element_text(face = "bold",
                                  hjust = 0.5)) 



allan <- allan_names |>
  pivot_wider(names_from = "Name",
              values_from = "Count",
              values_fill = 0) |>
  filter(Year == 2000) |>
  mutate(Total = Allan + Allen + Alan,
         across(.cols = Allen:Allan, .fns = ~.x/Total,
                .names = "{col}Prop"),
         OverallProp = Total/sum(Total)) |>
  select(Year, Sex, State, Allen, AllenProp, Alan, AlanProp, Allan, 
         AllanProp, Total, OverallProp) |>
  print()



allan |>
  knitr::kable(format = "html",
               col.names = c("Year", 
                             "Sex", 
                             "State", 
                             "Allen", 
                             "Allen Proportion in State", 
                             "Alan", 
                             "Alan Proportion in State", 
                             "Allan", 
                             "Allan Proportion in State", 
                             "Total", 
                             "Overall State Proportion in Year"),
               caption = "Summary of Frequency and Proportion of 'Allen', 'Allan, and 'Allan' in Each State") |>
  kableExtra::kable_classic(html_font = "Arial") |>
  kableExtra::kable_styling(bootstrap_options = "striped")



allan_ca_pa <- allan |>
  filter(State %in% c("CA", "PA")) |>
  print()



allan_ca_pa |>
  knitr::kable(format = "html",
               col.names = c("Year", 
                             "Sex", 
                             "State", 
                             "Allen", 
                             "Allen Proportion in State", 
                             "Alan", 
                             "Alan Proportion in State", 
                             "Allan", 
                             "Allan Proportion in State", 
                             "Total", 
                             "Overall State Proportion in Year"),
               caption = "Summary of Frequency and Proportion of 'Allen', 'Allan, and 'Allan' in California and Pennsylvania",
               digits = 3) |>
  kableExtra::kable_classic(html_font = "Cambria") |>
  kableExtra::kable_styling(bootstrap_options = "striped", 
                            position = "left")
```
